---
title: "Capital Punishment in the U.S."
author: "Dan Miller"
date: "`r format(Sys.time(), '%A, %B %d, %Y @ %I:%M %p')`"
output:
  html_document:
    theme: yeti
    highlight: textmate
    code_folding: hide
    toc: yes
    toc_depth: 4
  pdf_document:
    toc: yes
    toc_depth: '4'
subtitle: STT2860 Project 02 Fall 2022
---
<hr>

```{r globaloptions, include = FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  comment = NA
)
```


### Packages Used

```{r}
library(tidyverse)
library(readr)
```

### Read Data Files

```{r}
death_row_size <- readRDS("deathrowsize.rds")
death_sentences_NC <- readRDS("deathsentencesNC.rds")
cap_punish <- readRDS("cappunish.rds")
```

### Plot #1: How Many States Have Capital Punishment? 

* Create a barplot for the variable `dp2`. 
* Use `color =` and `fill =` to format the bars.
* Use `labs()` to add a title and better axis labels.
* Apply a built-in theme from **ggplot2** or **ggthemes**.
* You may optionally add a subtitle or caption to the plot.
* This plot should *not* have a visible legend anywhere on it.

```{r}
cap_punish %>%
ggplot(aes(x = dp2)) + 
  geom_bar(aes(fill = dp2), color = "black") +
  labs(title = "State Decisions on Capital Punishment", 
       x = "Decision", 
       y = "Number of States") +
  theme(plot.title = element_text(hjust = 0.5, size = 18),
        legend.position = "none")
```

### Plot #2: Capital Punishment By Region of the U.S.

* Create a barplot for the variable `region`. 
* Use `fill =` to shade each bar according to `dp2`.
* Change the `position =` to set each bar equal to 100%.
* Use `width =` to make the bars narrower than the default.
* Use `labs()` to add a title, axis labels, and a better legend label.
* Use `scale_fill_manual()` to choose a different color scheme for the plot. 
* Use `theme()` to move the legend to the bottom of the plot instead of the right
* You may optionally add a subtitle, caption, or ggplot2 built-in theme to the plot.

```{r}
colors <- c("#66c2a5","#fc8d62","#8da0cb")
cap_punish %>%
ggplot(aes(x = region)) +
  geom_bar(aes(fill = dp2), position = "fill", width = 0.65) +
  scale_fill_manual(values = colors) +
  scale_fill_discrete("Has Death Penalty") +
  labs(title = "Death Penalty by Major US Region",
       x = "US Region", 
       y = "Ratio of States") +
  theme(legend.position = "bottom",
        plot.title = element_text(hjust = 0.5, size = 18))
```

### Plot #3: Number of Prisoners on Death Row

* Create a barplot for the total number of prisoners (y) in each year (x).
* Fill the bars using the variable `decade` and outline them all in black.
* Set the width of all the bars so there is no visible gap in between them.
* Scale the x-axis to include a tick mark at each decade (1970, 1980, etc).
* Scale the y axis so that it is numbered 0, 500, 1000, 1500, 2000, etc.
* Use `labs()` to add or change the title, axis labels, and legend label. 
* Use `theme()` to do the following:
    * center your title
    * set the plot background to white or blank
    * set all axis tick marks to transparent or blank
    * set the major x-axis gridlines to transparent or blank
    * set the major y-axis gridlines to light gray
* You may optionally change the bar color scheme used in your plot

```{r}
death_row_size %>%
ggplot(aes(x = year, y = total)) +
  geom_col(aes(fill = decade), color = "black", width = 1) +
  scale_x_continuous(breaks = seq(1970, 2020, 10)) +
  scale_y_continuous(breaks = seq(0, 4000, by = 500)) +
  labs(title = "Number of Prisoners on Death Row", x = "Year", y = "Number of Prisoners") +
  scale_fill_discrete(name = "Decade") +
  theme(axis.ticks = element_blank(),
        panel.background = element_blank(),
        panel.grid.major.y = element_line(color = "grey"),
        panel.grid.major.x = element_blank(),
        plot.title = element_text(hjust = 0.5, size = 18))
  
```

### Plot #4: Number of Executions Since 1976

* Create a violin plot for the variable `post1976` (use `y = 0`)
* Add a boxplot and set the width so it fits inside the violin plot
* Fill each plot with a color of your choice (be colorblind-aware)
* Scale the x-axis of the plot so that it counts in increments of 25
* Get rid of all y-axis markings/gridlines using `scale_y_continuous()`
* Add a title, x-axis label, and a caption to identify the large outlier
* Add the built-in theme `theme_linedraw()` or `theme_minimal` to the plot

Hint: You can use the `NULL` vector in functions like `scale_y_continuous()`

```{r}
cap_punish %>%
ggplot(aes(x = post1976, y = 0)) +
  geom_violin(fill = "darkorchid3") +
  geom_boxplot(fill = "aquamarine1", width = .1) +
  scale_x_continuous(breaks = seq(0, 600, by = 25)) +
  scale_y_continuous(breaks = NULL) +
  labs(title = "Number of Executions Since 1976", 
       x = "Number of Executions",
       caption = "The outlier is Texas") +
  theme_linedraw() 
  
    
```

### Plot #5: Death Sentences in NC Since 1976

Create a time series plot of death sentences in North Carolina. Use three geometries on this plot: `geom_area()`, `geom_line()` and `geom_point()`. Use `geom_vline()` to add a vertical line for 2001. The significance of this year is that the General Assembly modified the death penalty statute in several ways, including giving prosecutors the discretion to *not* seek the death penalty in cases where it previously would have been mandatory. Format and label your plot in such a way that the information it conveys would be clear to the reader (i.e., the plot is reasonably stand-alone).

```{r}
death_sentences_NC %>%
ggplot(aes(x = year, y = sentences)) +
  geom_line() +
  geom_area(fill = "darkslategray") +
  geom_point() +
  geom_vline(xintercept = 2001, linetype = "dashed") +
  labs(x = "Year", y = "Number of Sentences") +
    theme_minimal()
```

### Information Conveyed by Your Plots

List two pieces of information you can learn or comparisons you can make from each plot above.

#### Plot #1

* Item 1: States seem split almost precisely down the middle when it comes to the death penalty.
* Item 2: There is a small subset of yes states (approximately 3) who have put a temporary suspension on executions (moratorium). 



#### Plot #2

* Item 1: The Northeast is the only US region with no outright/active yes votes. It is predominately no, with what I guess are one or two states with a temporary stay on executions. My guess is New Hampshire or Maine. 
* Item 2: Both The south and midwest are purely yes or no votes. No states with a moratorium. (the South also has far and away the most "yes" states). 

#### Plot #3

* Item 1: The number of prisoners on death row rises almost exponentially from the mid 1970's early 80's all the way until 2000/2001. 
* Item 2: Then a sharp yoy decline after 9/11, then a somewhat more gradual decline that brings us to today.  

#### Plot #4

* Item 1:Texas is far and away the obvious outlier with 575 executions since 1976.
* Item 2: The majority cluster of executions post 76 per state is in the range of 0 - 10. There are 6 other notable states in the 75-125 range aswell.  

#### Plot #5

* Item 1: In the year 2020 there were no new death sentences issued in North Carolina.
* Item 2: 1995 was the year with the most new Death Sentences, appx 34.  


<hr>

```{r}
sessionInfo()
```