---
title: "Final Presentation"
author: "Dan Miller"
date: "`r format(Sys.time(), '%B %d, %Y @ %I:%M %p')`"
output:
  ioslides_presentation: default
  slidy_presentation: default
---

```{r global_options, include = FALSE}
knitr::opts_chunk$set(
  echo = FALSE, # hide all the code
  message = FALSE,
  warning = FALSE,
  comment = NA
)
```
## Packages Used

I used tidyverse for data manipulation and plotting.

```{r, echo = TRUE}
library(tidyverse)
```

I need plotly to create pie charts.

```{r, echo = TRUE}
library(plotly)
```


## Ice Cream Dataset

I retrieved my dataset from [Kaggle](https://www.kaggle.com/datasets/tysonpo/ice-cream-dataset)

**Variables in the Dataset:**

* `brand`: Ben & Jerry's, Haagen Dazs, Talenti, Breyers
* `key`: Unique code for each ice cream product
* `name`: The name of the ice cream product
* `description`: Official description from respective company
* `rating`: Based on aggregated google reviews with a max score of 5  
* `rate_count`: Total occurrence of reviews (rating) on google
* `quality` : Based on subjective `rating` ranges. Great >= 4.5, Decent >= 3.75, Average >= 3
              Not Good >= 2.2 and Bad < 2.2

## 
```{r}
data <- read.csv("~/Rstudio Projects/Git Repo's/Project 3/Final Project/products (1).csv")
```


```{r}
data_edit <- data %>%
  select(!c(subhead, description)) %>%
  mutate(brand = case_when(brand == "bj"~"Ben & Jerry's", brand == "hd"~"Haagen Dazs", brand == "talenti"~"Talenti", brand == "breyers"~"Breyers")) %>%
  mutate(quality = rating, quality = case_when(quality >= 4.5~'Great', quality >= 3.75~"Decent", quality >= 3~"Average", quality >= 2.2~"Not Good", TRUE~"Bad")) %>%
  mutate(quality = as.factor(quality)) %>% 
  select(brand, flavor = name, rating, quality, ingredients) %>%
  arrange(quality)
```



```{r, fig.align = "center"}
data_edit %>%
 ggplot(aes(x = brand)) + 
  geom_bar(aes(fill = brand), show.legend = FALSE, alpha = 0.75) +
  labs(title = "Ice Cream Number of Flavors by Brand", y = "Count", x = NULL) +
  scale_y_continuous(breaks = seq(0, 100, by = 5)) +
  theme(panel.background = element_blank(),
        panel.grid.major.y = element_line(color = "grey", linewidth = .35, linetype = 1),
        plot.title = element_text(hjust = 0.5, size = 22))
  
```

##
```{r , fig.align = "center", fig.width = 7.5, fig}
data_edit %>% 
  ggplot(aes(x = brand, fill = factor(quality, levels=c("Great", "Decent", "Average", "Not Good", "Bad")))) +
  geom_bar(position = "stack", alpha = 0.75) + 
  labs(title = "Quality Comparison", y = "Count", x = "Brands", caption = "* based on google reviews (max 5). Great >= 4.5, Decent >= 3.75, Average >= 3, Not Good >= 2.2, Bad < 2.2 ") +
  scale_fill_discrete(name = "Quality") +
  scale_y_continuous(breaks = seq(0, 100, by = 5)) +
  theme(panel.background = element_blank(),
        panel.grid.major.y = element_line(color = "grey", size = .35, linetype = 1),
        panel.grid.minor.y = element_blank(),
        plot.title = element_text(hjust = 0.5, size = 22))
```

##
```{r, fig.align = "center"}
colors <- c("#66c2a5","#fc8d62","#8da0cb","#e78ac3","#a6d854")

pie_charts <- data_edit %>%
  group_by(quality) %>%
  summarize(counts = n(), percentage = n()/nrow(data_edit))

pie <- ggplot(data = pie_charts, aes(x = "", y = percentage, fill = factor(quality, levels = c("Great", "Decent", "Average", "Not Good", "Bad")))) +
  geom_col(color = "black") +
  coord_polar("y", start=0) +
  geom_text(aes(label = paste0(round(percentage*100), "%")), position = position_stack(vjust = 0.5)) +
  theme(panel.background = element_blank(), 
        axis.line = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        plot.title = element_text(hjust = 0.5, size = 22),
        legend.title = element_blank()) +
  labs(y = element_blank()) +
  ggtitle("Average Quality Distribution") +
  scale_fill_manual(values = colors)
pie

```


##
```{r, fig.align = "center"}
bj_pie_charts <- data_edit %>%
  filter(brand == "Ben & Jerry's")
bj_pie <- bj_pie_charts %>%
  group_by(quality) %>%
  summarize(counts = n(), percentage = n()/nrow(bj_pie_charts))

pie_2 <- ggplot(data = bj_pie, aes(x = "", y = percentage, fill =  factor(quality, levels = c("Great", "Decent", "Average", "Not Good", "Bad")))) +
  geom_col(color = "black") +
  coord_polar("y", start=0) +
  geom_text(aes(label = paste0(round(percentage*100), "%")), position = position_stack(vjust = 0.5)) +
  theme(panel.background = element_blank(), 
        axis.line = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
       plot.title = element_text(hjust = 0.5, size = 22),
        legend.title = element_blank()) +
  ggtitle("Ben & Jerry's") +
  labs(y = element_blank()) +
  scale_fill_manual(values = colors)
pie_2

```




##
```{r, fig.align = "center"}
hd_pie_charts <- data_edit %>%
  filter(brand == "Haagen Dazs")
hd_pie <- hd_pie_charts %>%
  group_by(quality) %>%
  summarize(counts = n(), percentage = n()/nrow(hd_pie_charts))

pie_3 <- ggplot(data = hd_pie, aes(x = "", y = percentage, fill =  factor(quality, levels = c("Great", "Decent", "Average", "Not Good", "Bad")))) +
  geom_col(color = "black") +
  coord_polar("y", start=0) +
  geom_text(aes(label = paste0(round(percentage*100), "%")), position = position_stack(vjust = 0.5)) +
  theme(panel.background = element_blank(), 
        axis.line = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        plot.title = element_text(hjust = 0.5, size = 22),
        legend.title = element_blank()) +
  ggtitle("Haagen Dazs") +
  labs(y = element_blank()) +
  scale_fill_manual(values = colors)
pie_3

```





##
```{r, fig.align = "center"}
br_pie_charts <- data_edit %>%
  filter(brand == "Breyers")
br_pie <- br_pie_charts %>%
  group_by(quality) %>%
  summarize(counts = n(), percentage = n()/nrow(br_pie_charts))

pie_4 <- ggplot(data = br_pie, aes(x = "", y = percentage, fill =  factor(quality, levels = c("Great", "Decent", "Average", "Not Good", "Bad")))) +
  geom_col(color = "black") +
  coord_polar("y", start=0) +
  geom_text(aes(label = paste0(round(percentage*100), "%")), position = position_stack(vjust = 0.5)) +
  theme(panel.background = element_blank(), 
        axis.line = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        plot.title = element_text(hjust = 0.5, size = 22),
        legend.title = element_blank()) +
  ggtitle("Breyers") +
  labs(y = element_blank()) +
  scale_fill_manual(values = colors)
pie_4

```



##
 
```{r, fig.align = "center"}
t_pie_charts <- data_edit %>%
  filter(brand == "Talenti")
t_pie <- t_pie_charts %>%
  group_by(quality) %>%
  summarize(counts = n(), percentage = n()/nrow(t_pie_charts))

pie_5 <- ggplot(data = t_pie, aes(x = "", y = percentage, fill =  factor(quality, levels = c("Great", "Decent", "Average", "Not Good", "Bad")))) +
  geom_col(color = "black") +
  coord_polar("y", start=0) +
  geom_text(aes(label = paste0(round(percentage*100), "%")), position = position_stack(vjust = 0.5)) +
  theme(panel.background = element_blank(), 
        axis.line = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        plot.title = element_text(hjust = 0.5, size = 22),
        legend.title = element_blank()) +
  ggtitle("Talenti") +
  labs(y = element_blank()) +
  scale_fill_manual(values = colors)
pie_5

```


##
* This content was not sponsored

